Co-citations in context

### 2019-05-31 cocit_r.tex
revised intro

% \begin{authorsummary} 
% \end{authorsummary}

\section{Introduction}

Citation and network analysis of  scientific literature reveals information on semantic relationships between publications, collaboration between scientists, and the practice of citation itself~\citep{garfield_citation_1955,de_solla_price_networks_1965,newman_structure_2001,Shi:2010:CHI:1816123.1816131,patience_pmid28560354}. Co-citation, the frequency with which two documents are cited together in other documents provides additional insights, including the identification of semantically related documents, fields, ideas, and specializations in science~\citep{small_co-citation_1973,marshakova-shaikevich_co-citation_1973,boyack_co-citation_2010, 10.3389/frma.2018.00020}.

\citep{uzzi_atypical_2013} used a novel approach for co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS)  to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. Observed co-citation frequencies, mapped to journals, were computed across the dataset by Monte Carlo simulation under a random graph model and normalized (shifted and scaled) to expected values. These normalized journal pair frequencies were termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores corresponding with the journal pairs represented in its citations. For each article, positional statistics of these z-scores were calculated to describe conventionality: high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the converse (LC). Similarly,  an article was deemed to have high novelty (HN) if the tenth percentile of its z-scores  was negative and low novelty (LN) for the converse. Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCHN (denoting that the article exhibits high conventionality and high novelty), with all four combinations being possible. 
\citep{uzzi_atypical_2013} observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a recipe for impact across the scientific literature.

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. A careful examination of their Monte Carlo simulation reveals that random substitutions of references used to generate expected journal pair frequencies are equiprobable, which means that a reference in an article can be substituted by any other reference randomly chosen from those published in the same year, irrespective of subject matter, disciplinary focus and citation count. 
For example, the model permits substitution of a reference in quantum physics with equal probability by a reference in quantum biology (i.e., the same field) or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is inconsistent with the actual frequency of citation and also with the power law or lognormal citation distributions observed in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  
Lastly,  as we will show here, while the citation switching algorithm used in Uzzi  {\em et al.} preserves the number of publications and the number of references in each publication, it does not preserve the proportion of cited references from within the disciplinary network.
 Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing journal pairs in terms of conventionality and novelty  can be reasonably questioned on grounds of model mis-specification.

A follow-up study by \citep{boyack_vs_uzzi_2014} explored the impact of discipline and journal effects on the definition of conventionality and novelty.  While their study had some methodological differences from \citep{uzzi_atypical_2013} in the use of Scopus data and a $\chi^2$ calculation for the expected number of journal pair citations, Boyack and Klavans find the same basic trend that HCHN is more probably in highly cited papers. However, they note that ``only 64.4\%  of  243  WoS  subject  categories'' in the Uzzi {\em et al.} study met the criterion of having the highest probability of hit papers in the HCHN category.  Further, they observed that journals vary widely in terms of size and influence and that 20 journals accounted for nearly 15\% of co-citations in their measurements. Lastly, they noted that three multidisciplinary journals accounted for 9.4\% of all atypical combinations, suggesting strong effects from both disciplines and journals that were not reported by Uzzi et al.

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In contrast, we chose to first construct semantically related sets of documents (disciplinary networks) and to measure observed and expected frequencies within these networks in order to preserve the proportion of citations from a citing document's subject area under random sampling.

We hypothesized that assessing articles' novelty, conventionality, and impact on datasets focused in disciplinary areas would reduce model mis-specification and better simulate citation practice, in turn, leading to different and more relevant conclusions. Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around academic disciplines. Within these disciplinary frameworks, we calculated expected journal pair frequencies using a random graph model that accounted for existing citation frequencies and an efficient Monte Carlo simulation algorithm that permitted a substantially greater number of simulations than do previous studies. We found our model did, in fact, preserve the citation patterns found in the literature whereas the approach of Uzzi et al. does not. The varying citation patterns when sampling from a disciplinary-focused dataset versus a broader dataset due, was a cause of journal pair z-scores changing signs in 28.6\% of instances: the effect of any one of these journal pairs on an article's novelty or conventiality is contradictory between broad and, more narrow disciplinary datasets. We contend that the interpretation due to a disciplinary dataset, which preserved citation patterns, is more appropriate and an improvement on current methods. Furthermore, our analysis of these semantically-themed citation networks challenges the universality of articles in the HCHN category having the highest probability of being hits as found in \cite{uzzi_atypical_2013}. For example, while HC remains highly correlated with hit articles in immunology and metabolism research, HN is not. In addition, while HN is highly correlated with hit articles in applied physics, HC is not.  Furthermore, we found that the categories demonstrating the highest percentage of hits is not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073).  For each of these years, we further restricted analysis to those of type Article. Since WoS data also contains incomplete references or references that point at other indexes,  we also considered only those references for which there were complete records (Table 1). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset. \\
\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005. Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched subject labels in WoS for `applied physics'. \\
%\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005. Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched article level labels in WoS for `applied physics'. \\
\emph{Normalization of observed and expected values} Building upon prior work~\cite{uzzi_atypical_2013}, all ${n \choose 2}$ reference pairs were generated for each publication, where $n$ is the number of cited references in the publication. These reference pairs were then mapped to the journals they were published in using ISSN numbers to create journal pairs. Where multiple ISSN numbers exist for a journal, the most frequently used one in the WoS was assigned to the journal. In addition, publications containing fewer than two references were discarded. Journal pair frequencies were summed across the dataset to create observed frequencies $(F_{obs})$. 
In contrast to the preceding study~\citep{uzzi_atypical_2013}, we generated 1,000 rather than 10 null models for each dataset by randomly shuffling references while preserving the number of publications, the number of references in each publication, and the frequency with which these references were cited within the year of interest. In contrast to the preceding study~\citep{uzzi_atypical_2013}, we generated 1,000 rather than 10 null models for each dataset.  Expected values $(F_{exp})$ for journal pairs were generated by averaging the result of 1,000 simulations. z-scores were calculated for each journal-pair using the formula $(F_{obs} - F_{exp})/\sigma$ where $\sigma$
 is the standard deviation of the frequencies generated by simulation. As a result of these calculations, each publication becomes associated with a set of z-scores corresponding to the journal pairs derived from pairwise combinations of its cited references and positional statistics (quantiles) of z-scores were calculated for each publication. Publications were subsequently labeled according to conventionality and novelty: (i) HC if the median z-score exceeded the median of median z-scores for all publications and LC if the median z-score was equal to or less than the median of median z-scores for all publications, 
and (ii) HN if the tenth percentile of z-scores for a publication was less than zero, and LN if the tenth percentile of z-scores for a publication was greater than zero. (Should we comment here that while we are using the term z-score, reusing it from Uzzi, we do not imply that these are related to a normal distribution?)

@@

% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of WoS Analytical Dataset. UP: unique publications, UR: unique references, TR: total references. The number of publications, unique references, total references and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Data for reference years is flanked by horizontal lines and shown in boldface. Only publications of type Article and references with complete WoS records are included in these counts.}
\label{tab:label}
\centering
\begin{tabular}{|rrrrr|}
  \hline
Year & UP & UR & TR & TR/UR \\ 
  \hline
\textbf{1985} & \textbf{418495} & \textbf{2281297} & \textbf{5615496} & \textbf{2.46} \\ 
\hline
1986 & 402309 & 2316451 & 5708796 & 2.46 \\ 
1987 & 412936 & 2427347 & 5998513 & 2.47 \\ 
1988 & 426001 & 2545647 & 6354917 & 2.50 \\ 
1989 & 443144 & 2673092 & 6749319 & 2.52 \\ 
1990 & 458768 & 2827517 & 7209413 & 2.55 \\ 
1991 & 477712 & 2977784 & 7729776 & 2.60 \\ 
1992 & 492181 & 3134109 & 8188940 & 2.61 \\ 
1993 & 504488 & 3278102 & 8676583 & 2.65 \\ 
1994 & 523660 & 3458072 & 9255748 & 2.68 \\ 
 \hline
\textbf{1995} & \textbf{559685} & \textbf{3692575} & \textbf{9897946} & \textbf{2.68} \\ 
\hline
1996 & 663110 & 4144581 & 11641286 & 2.81 \\ 
1997 & 677077 & 4340733 & 12135104 & 2.80 \\ 
1998 & 693531 & 4573584 & 12728629 & 2.78 \\ 
1999 & 709827 & 4784024 & 13280828 & 2.78 \\ 
 2000 & 721926 & 5008842 & 13810746 & 2.76 \\ 
 2001 & 727816 & 5203078 & 14261189 & 2.74 \\ 
 2002 & 747287 & 5464045 & 15001390 & 2.75 \\ 
 2003 & 786284 & 5773756 & 16024652 & 2.78 \\ 
 2004 & 826834 & 6095594 & 17167347 & 2.82 \\ 
 \hline
\textbf{2005} & \textbf{916573} & \textbf{6629595} & \textbf{19066249} & \textbf{2.88} \\ 
 \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Disciplinary Datasets. PubMed and WoS were searched for articles using search terms, `immunology', `metabolism', and `applied physics'. Counts of retrieved publications are shown for each of the three years analyzed.}
\label{tab:label}
\centering
\begin{tabular}{|r r r r|}
  \hline
Year & Immunology & Metabolism & Applied Physics \\ 
  \hline
1985 & 21,606 & 78,998 & 10,298 \\ 
1995 & 29,320 & 121,247 & 21,012  \\ 
2005 & 37,296 & 200,052 & 35,600  \\ 
 \hline
\end{tabular}
\end{table}

\section{Results}
\subsection{Chacko} In our study, we also use a Monte Carlo approach to simulate under a random graph model. A  principal consideration, however, was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by restricting random substitutions to only those references in the disciplinary network being studied. We estimated model misspecification by measuring the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies for the set of journals common to both a disciplinary network and the WoS superset (Table 3). The results indicate that simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset and its attendant substitutions that are ectopic with respect to field and discipline (Introduction).

\begin{table}[ht]
\caption{Measuring Model Misspecification. For the set of journal pairs in common between a disciplinary network and the full WoS dataset, Kullback-Leibler (K-L) divergences
between empirical and simulated journal pair frequencies were computed for the years 1985, 1995, and 2005 for the three disciplinary datasets (applied\_physics, immunology, and metabolism) using either the disciplinary network as background or the WoS superset (all\_wos) to generate the null model (Background). K-L divergence was calculated using the R seewave package with a base (logarithm) of 2. The ratio between the K-L divergence for disciplinary networks versus the full WoS ranges from 1.96 to 2.77 and is greater than 2.0 for eight out of nine cases, strongly suggesting that simulations that constrain substitutions to the given disciplinary network better model the observed data.}
\label{tab:label}
\centering
\begin{tabular}{|r lrlr r|}
  \hline
 & Disciplinary Network & Year & Background & K-L Divergence & Ratio \\ 
  \hline
1 & appl\_physics & 1985 & appl\_physics & 1.21 &  \\ 
  2 &  & 1985 & all\_wos & 2.37 & 1.96 \\ 
  3 &  & 1995 & appl\_physics & 0.86 &  \\ 
  4 &  & 1995 & all\_wos & 2.37 & 2.77 \\ 
  5 &  & 2005 & appl\_physics & 0.95 &  \\ 
  6 &  & 2005 & all\_wos & 2.35 & 2.47 \\ 
    \hline
  7 & immunology & 1985 & immunology & 0.75 &  \\ 
  8 &  & 1985 & all\_wos & 1.68 & 2.24 \\ 
  9 &  & 1995 & immunology & 0.78 &  \\ 
  10 &  & 1995 & all\_wos & 1.70 & 2.19 \\ 
  11 &  & 2005 & immunology & 0.73 &  \\ 
  12 &  & 2005 & all\_wos & 1.92 & 2.63 \\ 
    \hline
  13 & metabolism & 1985 & metabolism & 1.11 &  \\ 
  14 &  & 1985 & all\_wos & 2.24 & 2.02 \\ 
  15 &  & 1995 & metabolism & 1.07 &  \\ 
  16 &  & 1995 & all\_wos & 2.33 & 2.17 \\ 
  17 &  & 2005 & metabolism & 1.19 &  \\ 
  18 &  & 2005 & all\_wos & 2.60 & 2.18 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

% \begin{authorsummary} 
% \end{authorsummary}
\section{Introduction}

Citation and network analysis of  scientific literature reveals information on semantic relationships between publications, collaboration between scientists, and the practice of citation itself~\citep{garfield_citation_1955,de_solla_price_networks_1965,newman_structure_2001,Shi:2010:CHI:1816123.1816131,patience_pmid28560354}. Co-citation, the frequency with which two documents are cited together in other documents provides additional insights, including the identification of semantically related documents, fields, ideas, and specializations in science~\citep{small_co-citation_1973, marshakova-shaikevich_co-citation_1973,boyack_co-citation_2010, 10.3389/frma.2018.00020}.

\citep{uzzi_atypical_2013} used a novel approach for co-citation analysis of 17.9 million articles stratified by year of publication, and their cited references, from the Web of Science (WoS)  to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. In this study, frequencies of cited references in articles published in the Web of Science were calculated and transformed to journal pairs (observed co-citation frequencies). Expected values of these journal pair co-citations were computed from Monte Carlo simulations under a random graph model. Observed frequencies were then normalized (shifted and scaled) to expected values. These normalized journal pair frequencies were termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores corresponding with the journal pairs represented in its citations. For each article, positional statistics of these z-scores were calculated to describe conventionality: high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the converse (LC). Similarly,  an article was deemed to have high novelty (HN) if the tenth percentile of its z-scores  was negative and low novelty (LN) for the converse. Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCHN (denoting that the article exhibits high conventionality and high novelty), with all four combinations being possible. 
\citep{uzzi_atypical_2013} observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a recipe for impact across the scientific literature.

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. A careful examination of their Monte Carlo simulations reveals that random selection of the references used to generate a null model and expected values is equiprobable with respect to disciplinary origin and citation count.  For example, the model permits substitution of a reference in quantum physics with equal probability by a reference in the related quantum chemistry field or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is inconsistent with citation behavior and inconsiderate of the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  
Lastly,  as we will show, while the citation switching algorithm used in Uzzi  {\em et al.} preserves the number of publications and the number of references in each publication, it is not designed to preserve disciplinary proportions of cited references within subsets of the Web of Science. Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing journal pairs in terms of conventionality and novelty  can be reasonably questioned on grounds of model mis-specification.

A follow-up study by \citep{boyack_vs_uzzi_2014} explored the impact of discipline and journal effects on the definition of conventionality and novelty.  While their study had some methodological differences from \citep{uzzi_atypical_2013} in the use of Scopus data and a $\chi^2$ calculation rather than Monte Carlo simulations to generate expected values of journal pair citations, Boyack and Klavans find the same basic trend that HCHN is more probably in highly cited papers. However, they note that ``only 64.4\%  of  243  WoS  subject  categories'' in the Uzzi {\em et al.} study met the criterion of having the highest probability of hit papers in the HCHN category.  Further, they observed that journals vary widely in terms of size and influence and that 20 journals accounted for nearly 15\% of co-citations in their measurements. Lastly, they noted that three multidisciplinary journals accounted for 9.4\% of all atypical combinations, suggesting strong effects from both disciplines and journals that were not reported by Uzzi et al.

Despite different methods used to generate expected values, both \citep{uzzi_atypical_2013} and \citep{boyack_vs_uzzi_2014} measured observed co-citation frequencies across the scientific literature without disciplinary constraints and subsequently used normalized frequencies to examine disciplinary subsets. In extending this prior work, we hypothesized that analysis of datasets focused in disciplinary areas would reduce model mis-specification since the pool of citations would be relevant to those disciplinary areas rather than being drawn from all disciplines. Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around academic disciplines. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm. The null model generated by citation shuffling in our study preserves the number of publications, the number of references group by year within each publication and, importantly, the disciplinary proportions of cited references in each dataset.

Our analysis of these semantically-themed citation networks challenges the concept of universal features of the scientific literature that can be described by co-citation patterns. Particularly that articles in the HCHN category have the highest probability of being hits as found in \cite{uzzi_atypical_2013}. We find, for example, while HC remains highly correlated with hit articles in the immunology and metabolism datasets we constructed, HN is not. In addition, while HN is highly correlated with hit articles in applied physics, HC is not.  Furthermore, we found that the categories demonstrating the highest percentage of hits are not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

\emph{these remaining sentences require some work}
The varying citation patterns when sampling from a disciplinary-focused dataset versus a broader dataset due, was a cause of journal pair z-scores changing signs in 28.6\% of instances: the effect of any one of these journal pairs on an article's novelty or conventionality is contradictory between broad and, more narrow disciplinary datasets. We contend that the interpretation due to a disciplinary dataset, which preserved citation patterns, is more appropriate and an improvement on current methods.

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073).  For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(\ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset. \\
\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(\ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.

\emph{Normalization of observed and expected values} Building upon prior work~\cite{uzzi_atypical_2013}, all ${n \choose 2}$ reference pairs were generated for each publication, where $n$ is the number of cited references in the publication. These reference pairs were then mapped to the journals they were published in using ISSN numbers to create journal pairs. Where multiple ISSN numbers exist for a journal, the most frequently used one in the WoS was assigned to the journal. In addition, publications containing fewer than two references were discarded. Journal pair frequencies were summed across the dataset to create observed frequencies $(F_{obs})$. 
In contrast to the preceding study~\citep{uzzi_atypical_2013}, we generated 1,000 rather than 10 null models for each dataset by randomly shuffling references while preserving the number of publications, the number of references in each publication, and the frequency with which these references were cited within the year of interest. In contrast to the preceding study~\citep{uzzi_atypical_2013}, we generated 1,000 rather than 10 null models for each dataset.  Expected values $(F_{exp})$ for journal pairs were generated by averaging the result of 1,000 simulations. z-scores were calculated for each journal-pair using the formula $(F_{obs} - F_{exp})/\sigma$ where $\sigma$
 is the standard deviation of the frequencies generated by simulation. As a result of these calculations, each publication becomes associated with a set of z-scores corresponding to the journal pairs derived from pairwise combinations of its cited references and positional statistics (quantiles) of z-scores were calculated for each publication. Publications were subsequently labeled according to conventionality and novelty: (i) HC if the median z-score exceeded the median of median z-scores for all publications and LC if the median z-score was equal to or less than the median of median z-scores for all publications, 
and (ii) HN if the tenth percentile of z-scores for a publication was less than zero, and LN if the tenth percentile of z-scores for a publication was greater than zero. (Should we comment here that while we are using the term z-score, reusing it from Uzzi, we do not imply that these are related to a normal distribution?)

@@

% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of WoS Analytical Dataset. UP: unique publications, UR: unique references, TR: total references. The number of publications, unique references, total references and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Data for reference years is flanked by horizontal lines and shown in boldface. Only publications of type Article and references with complete WoS records are included in these counts.}
\label{tab:summary_data}
\centering
\begin{tabular}{|rrrrr|}
  \hline
Year & UP & UR & TR & TR/UR \\ 
  \hline
\textbf{1985} & \textbf{418495} & \textbf{2281297} & \textbf{5615496} & \textbf{2.46} \\ 
\hline
1986 & 402309 & 2316451 & 5708796 & 2.46 \\ 
1987 & 412936 & 2427347 & 5998513 & 2.47 \\ 
1988 & 426001 & 2545647 & 6354917 & 2.50 \\ 
1989 & 443144 & 2673092 & 6749319 & 2.52 \\ 
1990 & 458768 & 2827517 & 7209413 & 2.55 \\ 
1991 & 477712 & 2977784 & 7729776 & 2.60 \\ 
1992 & 492181 & 3134109 & 8188940 & 2.61 \\ 
1993 & 504488 & 3278102 & 8676583 & 2.65 \\ 
1994 & 523660 & 3458072 & 9255748 & 2.68 \\ 
 \hline
\textbf{1995} & \textbf{559685} & \textbf{3692575} & \textbf{9897946} & \textbf{2.68} \\ 
\hline
1996 & 663110 & 4144581 & 11641286 & 2.81 \\ 
1997 & 677077 & 4340733 & 12135104 & 2.80 \\ 
1998 & 693531 & 4573584 & 12728629 & 2.78 \\ 
1999 & 709827 & 4784024 & 13280828 & 2.78 \\ 
 2000 & 721926 & 5008842 & 13810746 & 2.76 \\ 
 2001 & 727816 & 5203078 & 14261189 & 2.74 \\ 
 2002 & 747287 & 5464045 & 15001390 & 2.75 \\ 
 2003 & 786284 & 5773756 & 16024652 & 2.78 \\ 
 2004 & 826834 & 6095594 & 17167347 & 2.82 \\ 
 \hline
\textbf{2005} & \textbf{916573} & \textbf{6629595} & \textbf{19066249} & \textbf{2.88} \\ 
 \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Disciplinary Datasets. PubMed and WoS were searched for articles using search terms, `immunology', `metabolism', and `applied physics'. Counts of retrieved publications are shown for each of the three years analyzed.}
\label{tab:disc-datasets}
\centering
\begin{tabular}{|r r r r|}
  \hline
Year & Immunology & Metabolism & Applied Physics \\ 
  \hline
1985 & 21,606 & 78,998 & 10,298 \\ 
1995 & 29,320 & 121,247 & 21,012  \\ 
2005 & 37,296 & 200,052 & 35,600  \\ 
 \hline
\end{tabular}
\end{table}

\section{Results}
\subsection{Chacko} In our study, we also use a Monte Carlo approach to simulate under a random graph model. A  principal consideration, however, was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied. We predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results indicate that for the set of journals common to both a disciplinary network and the WoS superset~\ref{tab:KLD}, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}

\begin{table}[ht]
\caption{Measuring Model Misspecification. For the set of journal pairs in common between a disciplinary network and the full WoS dataset, Kullback-Leibler (K-L) divergences
between empirical and simulated journal pair frequencies were computed for the years 1985, 1995, and 2005 for the three disciplinary datasets (applied\_physics, immunology, and metabolism) using either the disciplinary network as background or the WoS superset (all\_wos) to generate the null model (Background). K-L divergence was calculated using the R seewave package with a base (logarithm) of 2. The ratio between the K-L divergence for disciplinary networks versus the full WoS ranges from 1.96 to 2.77 and is greater than 2.0 for eight out of nine cases, strongly suggesting that simulations that constrain substitutions to the given disciplinary network better model the observed data.}
\label{tab:kld}
\centering
\begin{tabular}{|r lrlr r|}
  \hline
 & Disciplinary Network & Year & Background & K-L Divergence & Ratio \\ 
  \hline
1 & appl\_physics & 1985 & appl\_physics & 1.21 &  \\ 
  2 &  & 1985 & all\_wos & 2.37 & 1.96 \\ 
  3 &  & 1995 & appl\_physics & 0.86 &  \\ 
  4 &  & 1995 & all\_wos & 2.37 & 2.77 \\ 
  5 &  & 2005 & appl\_physics & 0.95 &  \\ 
  6 &  & 2005 & all\_wos & 2.35 & 2.47 \\ 
    \hline
  7 & immunology & 1985 & immunology & 0.75 &  \\ 
  8 &  & 1985 & all\_wos & 1.68 & 2.24 \\ 
  9 &  & 1995 & immunology & 0.78 &  \\ 
  10 &  & 1995 & all\_wos & 1.70 & 2.19 \\ 
  11 &  & 2005 & immunology & 0.73 &  \\ 
  12 &  & 2005 & all\_wos & 1.92 & 2.63 \\ 
    \hline
  13 & metabolism & 1985 & metabolism & 1.11 &  \\ 
  14 &  & 1985 & all\_wos & 2.24 & 2.02 \\ 
  15 &  & 1995 & metabolism & 1.07 &  \\ 
  16 &  & 1995 & all\_wos & 2.33 & 2.17 \\ 
  17 &  & 2005 & metabolism & 1.19 &  \\ 
  18 &  & 2005 & all\_wos & 2.60 & 2.18 \\ 
   \hline
\end{tabular}
\end{table}
